The Analogue Method for Precipitation Forecasting: Finding Better Analogue Situations at a Sub-Daily Time Step

### 2016-05-18 Moving temporal window - HESS/template.tex
Corrections from Michel.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogues. The constraint being the use of daily time series, due to the length of available archives they provide. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogues at a different hour, a sliding time window (STW) has been introduced.

@@

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analogue situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the S1 criteria \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window and at certain hours.

In fact, the predictand is not an instantaneous value, but often an accumulation over a certain time period (eg. 24 h for rainfall, taken from 6-30 h UTC). Therefore, the meteorological situation generating this accumulated rainfall is not instantaneous either, but is a continuous transition between changing circulation patterns. Thus, the observed predictors fields are considered at different time of the day, and at different heights. In our case, the geopotential height is considered at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The selection of the observation time of the predictor was found by \citet{Bontron2004} to have a significant influence. Moreover, due to inter-correlation, selecting the optimal 2 pressure levels at 2 hours is not equivalent to combining the best time of both levels individually.

@@

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{n} \sum_{i=1}^{n} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. The mean CRPS is averaged on the calibration, respectively the validation periods, on all days. \citet{Bontron2004} proposed a decomposition of this score interpreted into a sharpness and an accuracy terms.

@@

\begin{equation}
\label{eq:CRPSS}
CRPSS = \frac{CRPS-CRPS_{r}}{CRPS_{p}-CRPS_{r}} = 1-\frac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the CRPS value for the reference and $CRPS_{p}$ would be the one for a perfect forecast (which implies $CRPS_{p}~=~0$). A better forecast is characterized by an increase in CRPSS.

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analogue situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations do not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a sliding time window (STW) was introduced to allow searching for candidates at different hours of the day.

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (section \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving average (section \ref{sec:method}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Figure \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations.

@@

% FIGURE (MAP)

@@

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.3cm]{figures/illustration_moving_window.pdf}
	\end{center}
	\caption{Illustration of the principle of a sliding time window. }
	\label{fig:principle}
\end{figure}

>>>>

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day, for both the target day and the candidate analogues. The constraint being the use of daily time series, due to the length of available archives they provide, and the unavailability of equivalent archives at a finer time step. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogues at a different hour, a sliding time window (STW) has been introduced on a reduced archive of hourly precipitation totals.

@@

The first method is based on the analogy of the atmospheric circulation only \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. Searching analogue situations for a target day starts by a preselection step of the potential candidate for analogy. Here, it has been limited to the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the S1 criteria \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window and at certain hours:

The predictand, here a 24~h precipitation total, results from meteorological situations that are continuous transitions between changing circulation patterns. Thus, the observed predictors fields, extracted from reanalysis datasets, are considered at different time of the day and at different heights. In our case, the geopotential height is considered at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The selection of the observation time of the predictor was found by \citet{Bontron2004} to have a significant influence.

@@

\begin{equation}
\label{eq:CRPS}
CRPS = \frac{1}{n} \sum_{i=1}^{n} \left(  \int_{-\infty}^{+\infty} \left[ F_{i}(y)-H_{i}(y-y_{i}^{0})\right]^{2} dy \right) 
\end{equation}
where $H(y-y_{i}^{0})$ is the Heaviside function that is null when $y-y_{i}^{0}<0$, and has the value 1 otherwise. \citet{Bontron2004} proposed a decomposition of this score interpreted into a sharpness and an accuracy terms.

@@

\begin{equation}
\label{eq:CRPSS}
CRPSS = \frac{CRPS-CRPS_{r}}{CRPS_{p}-CRPS_{r}} = 1-\frac{CRPS}{CRPS_{r}}
\end{equation}
where $CRPS_{r}$ is the CRPS value for the reference (climatological distribution) and $CRPS_{p}$ would be the one for a perfect forecast (which implies $CRPS_{p}~=~0$). A better forecast is characterized by an increase in CRPSS.

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent at a finer resolution. Therefore, the analogue situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations does not occur systematically at the same time of the day, and that better candidates can be found by shifting to a different hour. On this assumption, a sliding time window (STW) was introduced to allow searching for candidates at different hours of the day.

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (section \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but starting at different 6-hourly time step, by means of a moving total (section \ref{sec:method}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Figure \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations. The precipitation time series were then aggregated by means of a moving 24~h-total (see next section).

@@

@@

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.3cm]{figures/illustration_moving_window.pdf}
	\end{center}
	\caption{Illustration of the principle of a sliding time window. The bars illustrate the 24~h precipitation totals.}
	\label{fig:principle}
\end{figure}
